Crowd density estimation from Wi-Fi positioning data in the Amsterdam Arena

### 2016-07-04 PaperOnSmoothing/draft-31052016.tex
typos corrected
Continued a bit on the Methods section.

We can assume that there is an upper limit $\lambda$ of a number of people that go to a concert together (a group), that is, whose locations are correlated\footnote{Actually, when the crowd is so dense that people can not move freely anymore, the whole crowd becomes a "group" (cite LP disaster) and all locations are correlated. But we aim to detect a high density with our method {\it before} this happens, to react preventively; otherwise, it is too late.}. Denote by $\kappa$ the maximal covariance between any $X_i$ and $X_j$ and let us write $i \sim j$ if and only if $i$ and $j$ are in the same group. 
Then, 
\begin{equation}\label{eq_covar}
\sum_{i \not =  j}Cov(X_i,X_j) = 2 \sum_{1\leq i <  j \leq n }Cov(X_i,X_j) =  2 \sum_{i,j: i \sim j }Cov(X_i,X_j)  \leq 2 n \cdot \frac{\lambda(\lambda - 1)}{2} \cdot\kappa = n  k \lambda(\lambda - 1)
\end{equation} 
where the inequality holds because the maximal number of groups is $n$ and the maximal number of pairs $(i,j)$ in a group is $\lambda(\lambda-1)/2$.
Let $\mu = \frac{1}{n}\sum_{i=1}^{n} Var(X_i)$, that is, 
denote by $\mu$ the average variance of $X_i$.  From \eqref{eq_Var} and \eqref{eq_covar} we have

\begin{equation}
Var(\frac{X}{n} ) \leq \frac{1}{n^2}(n \mu + n  k \lambda(\lambda - 1)) = \frac{1}{n}(\mu +  k \lambda(\lambda - 1)), 
\end{equation}
which tends to $0$ when $n \rightarrow \infty$. Note that we have greatly overestimated the variance with the inequation in \eqref{eq_covar}.

@@

\begin{equation}
\text{d}X(t)=\sqrt{2D}\text{d}W(t)
\end{equation}
where $W(t)$ is a Wiener process. The density function $f(t,x)$ of the displacement of $X$ in the time interval $[0,t]$ is known, and is a normal distribution with mean $X(0)$ and variance
\begin{equation}
\sigma^2=2Dt
\end{equation}
Thus, in case have no new observations, we can further smooth the density estimate from the previous epoch by convolution with a normal distribution. 
The variance is determined by the length $t$ of the time interval, and the (average) walking speed of pedestrians. The walking speed in turn can be related to the local crowd density via the so-called fundamental diagram.\\

>>>>

We can assume that there is an upper limit $\lambda$ of a number of people that go to a concert together (a group), that is, whose locations are correlated\footnote{Actually, when the crowd is so dense that people can not move freely anymore, the whole crowd becomes a "group" (cite LP disaster) and all locations are correlated. But we aim to detect a high density with our method {\it before} this happens, to react preventively; otherwise, it is too late.}. Denote by $\kappa$ the maximal covariance between any $X_i$ and $X_j$ and let us write $i \sim j$ if and only if $i$ and $j$ are in the same group. 
Then, 
\begin{equation}\label{eq_covar}
\sum_{i \not =  j}Cov(X_i,X_j) = 2 \sum_{1\leq i <  j \leq n }Cov(X_i,X_j) =  2 \sum_{i,j: i \sim j }Cov(X_i,X_j)  \leq 2 n \cdot \frac{\lambda(\lambda - 1)}{2} \cdot\kappa = n  \kappa \lambda(\lambda - 1)
\end{equation} 
where the inequality holds because the maximal number of groups is $n$ and the maximal number of pairs $(i,j)$ in a group is $\lambda(\lambda-1)/2$.
Let $\mu = \frac{1}{n}\sum_{i=1}^{n} Var(X_i)$, that is, 
denote by $\mu$ the average variance of $X_i$.  From \eqref{eq_Var} and \eqref{eq_covar} we have

\begin{equation}
Var(\frac{X}{n} ) \leq \frac{1}{n^2}(n \mu + n  \kappa \lambda(\lambda - 1)) = \frac{1}{n}(\mu +  \kappa \lambda(\lambda - 1)), 
\end{equation}
which tends to $0$ when $n \rightarrow \infty$. Note that we have greatly overestimated the variance with the inequation in \eqref{eq_covar}.

@@

\begin{equation}
\text{d}X(t)=\sigma\text{d}W(t)
\end{equation}
where $W(t)$ is a Wiener process. The density function $f(t,x)$ of the displacement of $X$ in the time interval $[0,t]$ is known, and is a normal distribution with mean $X(0)$ and variance
\begin{equation}
\text{Var(\textit{X(t)})}=\sigma^{2}t
\end{equation}
Thus, in case we have no new observations, we smooth the density estimate from the previous epoch by convolution with the normal density function. 
The variance is determined by the length $t$ of the time interval, and the walking speed of pedestrians. 
[Optional: We relate $\sigma$ to the walking speed of pedestrians by setting $\sigma^2=2Dt$, where ]
\begin{equation}
D=\frac{(\Delta X)^2}{2\Delta t}
\end{equation}
The walking speed can be related to the local crowd density via the so-called \textit{Fundamental Diagram}. 
When the density increases the movement of pedestrians becomes more impeded.
The fundamental diagram has been the subject of extensive study [examples].
Here we use the fundamental diagram as provided by ... [?]

Smoothing a two-dimensional (input) function by convolution with a Gaussian weight function that adapts its width to the local values of the input function is [complicated and] time-consuming.\\
